---
title: "Week 12 Lab"
author: 'Hanoaf Al Nasser: A18008237'
date: "11/07/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Section 1. Proportion og G/G in a population 

Downloaded a csv file from Ensemble < 
https://useast.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=959672880#373531_tablePanel

Here we read this csv file 


Find the proportion 

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```
```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

Now let's look at a different population. I picked the GBR. 

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

Find proportion of G|G

```{r}
round(table(gbr$Genotype..forward.strand.) / nrow(gbr) * 100, 2)
```


This variant is associated with childhood asthma is more frequent in the GBR population than the MKL population. 

Lets now dig into this further.

## Section 4: Population analysis 

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.


```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)

```

```{r}
nrow(expr)
```

```{r}
table(expr$geno)
```

```{r}
library(ggplot2)

```
 
 
Lets make a boxplot 

```{r}
ggplot(expr) + aes(geno, exp, fill = geno) + geom_boxplot(notch = TRUE)
```


> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

They have different distributions, G/G is significantly different from the A/G and the A/A. The G/G is associated with having a reduced expression of ORMDL3. 











